---
title: "Data Exploration"
author: "Sarina Singh Khaira"
date: "09/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)
```


```{r}
seasearch <- read_csv(here("clean_data/seasearch.csv"))
```

# Number of surveys each year

```{r}
seasearch %>%
  select(start_date_year, survey_key) %>%
  unique() %>%
  ggplot() +
  aes(x = start_date_year) +
  geom_bar()
```
Reliable data starts around 2004

# Number of times species are recorded
```{r}
seasearch_cleaned %>%
  filter(taxon_rank == "species") %>%
  group_by(scientific_name) %>%
  summarise(count = n()) %>%
  slice_max(count, n = 15) %>%
  ggplot(aes(x = fct_reorder(scientific_name, -count), y = count)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Most commonly sighted organisms are asterias rubens and echinus esculentus

# Percentage of surveys which saw an urchin or starfish per year

```{r}
seasearch %>% 
  group_by(survey_key, start_date_year) %>%
  summarise(starfish = scientific_name == "Asterias rubens",
            urchin = scientific_name == "Echinus esculentus") %>%
  unique() %>%
  group_by(start_date_year) %>%
  summarise(starfish_present_prop = mean(starfish == TRUE),
            urchin_present_prop = mean(urchin == TRUE)) %>%
  ggplot() +
  aes(x = start_date_year) +
  geom_line(aes(y = starfish_present_prop)) +
  geom_line(aes(y = urchin_present_prop), colour = "blue")
```
Perhaps there is also seasonality, how many surveys are completed in the winter?

```{r}
seasearch %>%
  select(start_date_month, survey_key) %>%
  unique() %>%
  ggplot() +
  aes(x = start_date_month) +
  geom_bar()
```

Number of surveys containing Asterias rubens and echinus esculentus per year

```{r}
seasearch %>%
  filter(scientific_name %in% c("Asterias rubens", "Echinus esculentus"),
         start_date_year >= 2004 ) %>%
  ggplot() +
  aes(x = start_date_year, fill = scientific_name) +
  geom_bar(position = "dodge")
```

